\documentclass[12pt]{article}
\usepackage[margin=1cm]{geometry}
\usepackage{enumitem}
\usepackage[]{hyperref}
\usepackage{libertine}
%\renewcommand{\section}[2]%
%        {\pagebreak[2]\vspace{1\baselineskip}%
%         \phantomsection\addcontentsline{toc}{section}{#1}%
%         \hspace{0in}%
%         \marginpar{
%         \raggedright \scshape #1}#2}
%\renewcommand{\section}[1]{ \noindent\fbox{ \parbox[t]{1in}{\scshape#1} } }
\renewcommand{\section}[1]{ \noindent\parbox[t]{1in}{\bfseries\scshape#1}} 

%\newcommand{\entry}[1]{ \noindent\fbox{ \parbox[t]{6in}{#1}} \vspace{0.1in}\\}
\newcommand{\entry}[1]{\noindent\parbox[t]{6in}{#1}\vspace{0.1in}\\}
\setlength\itemindent{0in}%
\begin{document}
{\huge ANH DO}\\
\noindent\rule{\textwidth}{1pt}

\section{Objective}
\entry{
My interest lies in multi-armed bandit problem and sequential decision making and 
I enjoy not only designing new algorithms but also proving the effectiveness of it. 
However, reinforcement learning suffer inefficient sampling which need a lot of samples during its training.
Therefore, I am interest in designing a provably efficient algorithm to overcome this weakness of reinforcement learning. 
}
\section{Contact}
\entry{
    \textit{Phone}: %(+84)976525773 
    \\
    \textit{Email}: \href{mailto:anhddo93@gmail.com}{anhddo93@gmail.com} \\
    \textit{Homepage}: \href{https://sites.google.com/view/anhddo}{https://sites.google.com/view/anhddo} 
}

\section{Education}
\entry{
    \textbf{John von Neumann Institute - Vietnam National University HCM}.\hfill 2018-2020\\
    M.S in Information and Communication Technologies. \\
    \textit{Thesis: "Regret Bounds for Optimistic Value Iteration in Average-Reward MDPs".}\\
    \\
    \textbf{Ho Chi Minh University of Technology - Vietnam National University HCM.}\hfill 2011-2016\\
    B.Eng, Computer Science. (Honor Class)\\
    \textit{Thesis: "Density-based object detection".}\\
}
\section{Publication}
\entry{
Regret Bounds for Optimistic Value Iteration in Average-Reward MDPs.(2020)\\
Yasin Abbasi-Yadkori, \textbf{Anh Do}, Botao Hao, Nevena Lazic, Csaba Szepesvari.
}
\section{Experience}
\entry{
AI Research Resident, \href{https://www.vinai.io}{\textbf{VinAI Research}}, Hanoi, Vietnam.\hfill Sep 2019\\
\if0
Topic: \textit{"Regret bounds for Optimistic Value Iteration in Average-Reward MDPs".}\\
\fi
\textbf{Supervisor}: 
\href{https://yasin-abbasi.github.io/}{Yasin Abbasi-Yadkori}\\
\\
Machine learning engineer, \href{https://hiip.asia/?lang=en}{\textbf{HIIP}}, Ho Chi Minh, Vietnam. \hfill May 2018 - Sep 2018 \\
Responsibility: Building sentiment analysis systems for an influencer flatform.
\if0
Sofware Engineer, \href{https://rollingant.com}{Rolling Ant}, Ho Chi Minh, Vietnam.\hfill    Jul 2016 - Sep 2017\\
Responsibility: Develop interactive 3D applications on mobile devices
\fi
}

\section{\vspace{0in}Awards}
\entry{
    \begin{itemize}[noitemsep,topsep=0pt, leftmargin=*]
        \item 3rd runner up Hackathon Vietnam AI Grand Challenge - Southern Region. \hfill 2019
            \\ \href{https://www.aigrandchallenge.com}{https://www.aigrandchallenge.com}.
        \item First prize hackathon John von Neumann Institute. \hfill 2019
        \item First prize winner of Entropy - Data Analytics Competition. \hfill 2018
            \\ \href{http://entropy.jvn.edu.vn }{http://entropy.jvn.edu.vn }
        \item Full-scholarship for John von Neumann Institute Master Program. \hfill 2018
        \item SAI Digital scholarship. In recognition of outstanding academic achievement. \hfill 2015
        \item VNG scholarship. \hfill 2014
        \item Gameloft scholarship.  \hfill 2013 \\
                In honor of an excellent academic performance at Ho Chi Minh University of Technology.
    \end{itemize}
}
\end{document}
